### 2021-02-18 Notes.Rmd
back up data exploration

---
title: "Notes"
output: html_notebook
author: Juan Rocha
---

To-do list:

1. Compute the import / export ratio per country
2. Same at commodity level: are there commodities that are purely trade driven?
3. Explore Paula's file: rank countries per number of companies on high risk
4. Extract main commodities
5. Characterize missing values: where are they missing (countries) and what vars are not very useful?
6. Construct networks between countries and corporations weighted by subsidiaries?