### 2021-06-10 networks.Rmd
visualizing latest data cleaned by Paula

---
title: "Networks of financial actors"
author: "Juan Rocha"
date: "6/10/2021"
output: 
  html_notebook:
      theme:
          code_font:
            google: Fira Code
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE, message=FALSE, warning = FALSE, error = FALSE,
    fig.width = 6, fig.height = 5, tidy = TRUE, dpi  = 300)
```

```{r}
library(tidyverse)
library(plotly)
library(patchwork)
library(network)

theme_set(theme_light(base_size = 10))
```

Read data:

```{r message=FALSE, warning=FALSE, error=FALSE}
df1 <- readxl::read_xlsx(
    path = "data/Paulas_files/001-Drivers_Increasing_Risk_ED.xlsx",
    sheet = 2,
    na = "NA") %>% 
    janitor::clean_names()

```

Currently the data set has `r df1 %>% nrow()` companies, in `r df1$country %>% unique() %>% length()` countries, dealing with `r df1$commodity %>% unique() %>% length()` commodities (including `NA`).

The data is very incomplete with respect to country locations

```{r}
df1 %>% 
    select(company, origin, country) %>% 
    skimr::skim()
```

```{r}
library(sna)
net_countries <-  df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% 
    # unique() %>% 
    count(name = "channel") %>% # a channel is acountry * commodities
    filter(!is.na(origin), !is.na(country)) %>% 
    network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = FALSE)

net_countries %v% "indegree" <- sna::degree(
    net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, cmode = "indegree")
net_countries %v% "outdegree" <- sna::degree(
    net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, cmode = "outdegree")
net_countries %v% "degree" <- sna::degree(net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE)

gplot(
    net_countries,
    vertex.col = alpha("purple", 0.25),
    label = network.vertex.names(net_countries),
    label.cex = 0.5, label.col = "grey50",
    vertex.border = 0, 
    vertex.cex = sqrt(0.2 + net_countries %v% "outdegree"),
    label.pos = 5,
    edge.col = alpha("grey50", 0.1),
    edge.lwd = log1p(net_countries %e% "channel"),
    displayisolates = TRUE, pad = 0.5,
    mode = "mds"
    )

# networkD3::simpleNetwork(
#     df1 %>%
#         select(company, origin, country) %>%
#         group_by(origin, country) %>%
#         # unique() %>%
#         count(name = "channel") %>% 
#         filter(!is.na(origin), !is.na(country)) ,
#     Source = "origin", Target = "country", zoom = TRUE,
#     linkColour = "grey", nodeColour = "blue")

```
Maybe fit a PCA to the data with network stats, commodities and so forth per country, and use the coordinates for visualization. Currently using a multi-dimensional scaling projection. Another way of "seeing" the network is by ranking which countries are the most influencial given their links to other countries (_outdegree_), once self-links are removed (companies whose country of origin and operation is the same):

```{r}
df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% 
    # unique() %>% 
    count(name = "channel") %>% # a channel is acountry * commodities
    filter(!is.na(origin), !is.na(country)) %>% 
    filter(origin != country) %>% 
    arrange(desc(channel))
```

The dataset has `r unique(df1$company) %>% length()` unique companies and `r unique(df1$commodity) %>% length()` commodities. Below a table that summarizes in how many different places the same company extract the same resource:

```{r}
df1 %>% 
    select(company, commodity) %>% 
    group_by(company, commodity) %>% 
    count(name = "geographies") %>% 
    arrange(desc(geographies))
```

```{r}
df1 %>% 
    select(company, commodity) %>% 
    group_by(company, commodity) %>% 
    count(name = "geographies") %>% 
    arrange(desc(geographies)) %>% 
    ggplot(aes(company, commodity)) +
    geom_tile(aes(fill = geographies)) +
    scale_fill_viridis_c() +
    theme_light(base_size = 6) +
    theme(axis.text.x = element_text(angle = 90)) 
```