### 2022-04-04 Notes.Rmd
some figs updates

---
title: "Notes"
output: html_notebook
author: Juan Rocha
---

To-do list:

1. Compute the import / export ratio per country
2. Same at commodity level: are there commodities that are purely trade driven?
3. ~~Explore Paula's file: rank countries per number of companies on high risk~~
4. ~~Extract main commodities~~
5. Characterize missing values: where are they missing (countries) and what vars are not very useful?
6. ~~Construct networks between countries and corporations weighted by subsidiaries?~~

Exponential random graph models:
- Include world governance indicators
- Inequality
- Triangles? Feedforward triangles -> theory on instability?
- Valued ergm with number of companes count on link weight?
- Create network of investors combined for all cases
- Create investor - country (where investors are based)
- The later two should provide descriptive stats.

Experiment with the HH-Index: https://en.wikipedia.org/wiki/Herfindahl–Hirschman_index

Cannot make it bipartite because:
The following vertices appear in both columns:
	- Bnp Paribas
	- Oji Holdings Corporation