### 2022-10-13 ocean_disclosure_project.Rmd
quick check of companies for fisheries study

---
title: "Exploring fishing data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Ocean disclosure project

```{r}
library(tidyverse)

# ocean disclosure project:
odp <- readxl::read_excel("data/005-Ocean_disclosure_database.xlsx", sheet = 2) |> 
    janitor::clean_names()
```

Summary:

```{r}
odp |> skimr::skim()
```
Data exist for the years `r odp$reporting_year |> as.numeric() |> range(na.rm = TRUE)`. It covers the following species <font color = "blue">`r odp$species |> unique()` </font> captured in:

```{r}
odp |> pull(catching_area) |> unique()
```

A preliminary list of companies fishing in "Gulf of Alaska":

```{r}
odp |> 
    filter(str_detect(catching_area, "Gulf of Alaska")) |> 
    pull(company) |> unique()
```

A preliminary list of companies fishing anchoveta in South America:

```{r}
odp |> 
    filter(species %in% c("Anchoveta", "Pacific anchoveta")) |> 
    pull(company) |> unique()
```

A preliminary list of companies fishing tuna:

```{r}
odp |> 
    filter(str_detect(species, "tuna")) |> 
    pull(company) |> unique()
```

Although not all of them are based in the Eastern Pacific.

```{r}
odp |> 
    filter(str_detect(species, "tuna")) 
```

```{r}
odp
```

## Peruvian anchoveta and sardine

```{r}
anch <- readxl::read_excel(
    "data/16_cuota_anchoveta-sardina_comun_v-x_2022_v20220908.xlsx",
    sheet = 2, range = "D6:L33") |> janitor::clean_names()
sard <- readxl::read_excel(
    "data/16_cuota_anchoveta-sardina_comun_v-x_2022_v20220908.xlsx",
    sheet = 2, range = "D39:L64") |> janitor::clean_names()

anch

```

```{r}
sard
```
Sardina española: It can be a mixture of companies, individual people, and coops

```{r}
sardesp <- readxl::read_excel("data/14_cuota_anchoveta-sardina_espanola_xv-iv_2022_v20220908.xlsx", sheet = "Publicacion web") |> 
    janitor::clean_names()

sardesp |> 
    select(organizacion_titular_area) |> 
    unique() |> 
    filter(!str_detect(organizacion_titular_area, "TOTAL ASIGNATARIO"))
```
Industrial actors only:

```{r}
sardesp <- readxl::read_excel("data/14_cuota_anchoveta-sardina_espanola_xv-iv_2022_v20220908.xlsx", sheet = 6, range = "B10:P47") |> 
    janitor::clean_names()

sardesp |> 
    select(titular_de_cuota_ltp) |> 
    filter(!is.na(titular_de_cuota_ltp)) |> 
    unique() 
```

## Alaskan pollock

```{r}
apo <- readxl::read_excel("data/002-Allocation_quotas_2020.xlsx", sheet = 2) |> 
    janitor::clean_names()

apo |> select(permit_holder) |> unique()
```

## Pacific tuna

Nothing on the folder

## Feedback from team

- ODP is a voluntary database, not complete. Missing Japan, China, etc.
- Clean up retailers? base on suppliers
- check Peter pan... data for pollock incomplete
- expect a lot of companies to be private
- skip orbis this time and instead of equity looking at debt
- Explore the database in the library...

## Next steps

1. Juan puts together a list of companies
2. Paula checks costs of licences at Library
3. Learning session with Paula. 
4. Check pollock info in Alaska (first NOAA, and Kevin)
5. Publish in Science

Global fishing watch:
 - IMO unique identifier of vessels: track over time
 - Paper Carmine 2020 OneEarth ? (Only for Tuna?) Check with Frida: how much we can trace back companies, and how much do we miss?