### 2022-12-22 Notes.Rmd
back ups: graphs for paper, some notes, and data collection script

---
title: "Notes"
output: html_notebook
author: Juan Rocha
---

To-do list:

1. Compute the import / export ratio per country
2. Same at commodity level: are there commodities that are purely trade driven?
3. ~~Explore Paula's file: rank countries per number of companies on high risk~~
4. ~~Extract main commodities~~
5. Characterize missing values: where are they missing (countries) and what vars are not very useful?
6. ~~Construct networks between countries and corporations weighted by subsidiaries?~~

Exponential random graph models:
- Include world governance indicators
- Inequality
- Triangles? Feedforward triangles -> theory on instability?
- Valued ergm with number of companes count on link weight?
- Create network of investors combined for all cases
- Create investor - country (where investors are based)
- The later two should provide descriptive stats.

Experiment with the HH-Index: https://en.wikipedia.org/wiki/Herfindahl–Hirschman_index

Cannot make it bipartite because:
The following vertices appear in both columns:
	- Bnp Paribas
	- Oji Holdings Corporation

One could force it to be bipartite by changing name according to category. Once bipartite, use the one mode projections to calculate network statistics more relevant for influence. For example a weighted betweenness or something like that.

### EGRMs

- use constraints on degree: keep degrees constant. But not sure how to code it
- Play with control features:
        - MCMCLE.maxit = < 60 which is the current default (how do I get the defaults?)

## Marine study

Title: Identifying companies and financial actors exposed to marine tipping points