### 2024-02-27 paper/manuscript.qmd

---
title: "Identifying financial actors exposed to marine tipping points"
format: pdf
editor: visual
date: "2024-02-27"
author: 
  - name: Juan C. Rocha
    orcid: 0000-0003-2322-5459
    email: juan.rocha@su.se
    affiliation:
      - name: Stockholm Resilience Centre, Stockholm University
abstract: >
  Here is the abstract
  multiple lines
keywords:
  - Tipping points
  - Financial actors
  - Regime shifts
---

## Outline

-   Introduction \[1000w\]

-   Method \[800w\]

-   Results \[1000w\]

-   Discussion \[800w\]

-   Conclusion \[400w\]

## Quarto

Quarto enables you to weave together content and executable code into a finished document.
To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code.
You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).